---
title: "Airline Final"
output: html_document
date: "2025-11-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(randomForest)
library(caret)
```

```{r}
df <- read_csv("ecommerce_customer_behavior_dataset.csv")
```

# Link to data: https://www.kaggle.com/datasets/umuttuygurr/e-commerce-customer-behavior-and-sales-analysis-tr?select=ecommerce_customer_behavior_dataset.csv 

```{r}
colSums(is.na(df))
```

```{r}
ggplot(df, aes(Customer_Rating)) +
  geom_histogram()
```
# Remove ID columns

```{r}
df <- df %>%
  select(-Order_ID, -Customer_ID)
```


# Just grab the months from the date column and then turn that into seasons. 

```{r}
library(lubridate)
df$Season <- case_when(
  month(df$Date) %in% c(12, 1, 2) ~ "Winter",
  month(df$Date) %in% c(3, 4, 5) ~ "Spring",
  month(df$Date) %in% c(6, 7, 8) ~ "Summer",
  month(df$Date) %in% c(9, 10, 11) ~ "Fall",
)
```

# Remove date column

```{r}
df <- df %>%
  select(-Date)
```

```{r}
summary(df)
```
```{r}
ggplot(df, aes(Age)) +
  geom_histogram()
```

# Check and make sure the category columns have a reasonable amount of choices, if not then create subsections

```{r}
n_distinct(df$Gender)
```
```{r}
unique(df$Gender)
```
```{r}
n_distinct(df$City)
```
```{r}
n_distinct(df$Product_Category)
```
```{r}
n_distinct(df$Payment_Method)
```
```{r}
n_distinct(df$Device_Type)
```
```{r}
ggplot(df, aes(Unit_Price)) +
  geom_histogram()
```

# Since Unit Price, and Total Price are very similar and in some cases the same I will drop unit price. I will also convert discount amount into a more simple yes or no discount column. 

```{r}
df <- df %>%
  mutate(Discount = ifelse(Discount_Amount > 0, "Yes", "No")) %>%
  select(-Unit_Price, -Discount_Amount) 
```


```{r}
df %>%
  count(Discount)
```

```{r}
str(df)
```

```{r}
df <- df %>%
  mutate(across(where(is.character), as.factor))
```

```{r}
str(df)
```
```{r}
df <- df %>%
  mutate(Is_Returning_Customer = as.factor(Is_Returning_Customer))
```

```{r}
summary(df)
```
```{r}
ggplot(df, aes(Total_Amount)) + 
  geom_histogram()
```

# Doing a log transformation of Total_Amount as it has a large right skew as some purchases were very large compared to others. 

```{r}
df$Log_Total_Amount <- log(df$Total_Amount)
```

```{r}
ggplot(df, aes(Log_Total_Amount)) + 
  geom_histogram()
```

# Did a log transformation on total amount since it has a very severe right skew from some much more expensive items. Going to use the logged version in the modeling upcoming. 

```{r}
df <- df %>%
  select(-Total_Amount)
```


```{r}
library(caret)
library(randomForest)

set.seed(8)
train_method_used = trainControl(method="cv", number = 5)

fit_caret_lm <- train(Customer_Rating ~ .,
                      data = df,
                      method = "lm",
                      trControl = train_method_used,
                      metric = "MAE")

fit_caret_rf = train(Customer_Rating ~ ., 
             data = df,
             method = "rf",
             trControl = train_method_used,
             tuneGrid = expand.grid(mtry = c(2,4,6)),
             ntree = 200,
             metric = "MAE")

all_MAE_CV <- c(min(fit_caret_lm$results$MAE),
                min(fit_caret_rf$results$MAE))

all_train_output <- list(fit_caret_lm,fit_caret_rf)

bestmodel_train_output <- all_train_output[[which.min(all_MAE_CV)]]

bestmodel_train_output
```
```{r}
fit_caret_lm
```

```{r}
fit_caret_rf$finalModel
```

# Random forest proved to be the best, with 2 variables tried at each split. The models are still not very good at predicting though, it has a negative percent Var explained, showing that it may be preforming worse than a simple model. We also see that the R squared is almost at 0, which shows that there is not much of the variation of customer reviews explained by this model, and that there is not a strong relationship between the response and predictor variables. 


```{r}
n = nrow(df)  
nfolds_outer = 5  
groups = rep(1:nfolds_outer,length=n)  
set.seed(8)
cvgroups = sample(groups,n)  

# set up storage for outer 5-fold cross-validation for model ASSESSMENT
allbesttrain = list(rep(NA,nfolds_outer))  
allpredicted_outer = rep(NA,n)  

train_method_used = trainControl(method="cv", number = 5)

for (jj in 1: nfolds_outer) {    
  in_train_outer = (cvgroups != jj)   
  in_test_outer = (cvgroups == jj)     

  train_set_outer = df[in_train_outer, ] 
  test_set_outer = df[in_test_outer, ]   
  
  dataused=train_set_outer
  
  
  fit_caret_lm_2 <- train(Customer_Rating ~ .,
                      data = dataused,
                      method = "lm",
                      trControl = train_method_used,
                      metric = "MAE")

  fit_caret_rf_2 = train(Customer_Rating ~ ., 
                     data = dataused,
                     method = "rf",
                     trControl = train_method_used,
                     tuneGrid = expand.grid(mtry = c(2,4,6)),
                     ntree = 200,
                     metric = "MAE")
        
  all_MAE_CV <- c(min(fit_caret_lm_2$results$MAE),
                  min(fit_caret_rf_2$results$MAE))
  
  all_train_output <- list(fit_caret_lm_2, fit_caret_rf_2)

  bestmodel_train_output <- all_train_output[[which.min(all_MAE_CV)]]




  allbesttrain[[jj]] = bestmodel_train_output

  allpredicted_outer[in_test_outer]  = bestmodel_train_output |> predict(test_set_outer)
}

# assessment
y = df$Customer_Rating
gf_point(y ~ allpredicted_outer)
```

```{r}
# This line will compute the models best estimate for future performance
MAE_outer = mean(abs(y - allpredicted_outer)); MAE_outer
```

# We see that we will likely be .882 rating off of future customer reviews. It is not terrible, but it is also not great. 

```{r}
min(fit_caret_lm_2$results$MAE)
min(fit_caret_rf_2$results$MAE)
```
# Random Forest has the lower MAE value

```{r}
final_model <- fit_caret_rf$finalModel
final_model
```
```{r, fig.height=6.5}
varImpPlot(final_model)
```
```{r}
varImp(final_model) %>%
  arrange(desc(Overall))
```

# We can see here which variables the model found most important. 


```{r}
cor(df$Customer_Rating, df$Log_Total_Amount)
cor(df$Customer_Rating, df$Session_Duration_Minutes)
cor(df$Customer_Rating, df$Age)
```
# As expected from the model results, the correlation of the response to even the 3 most important predictors is very small. 


```{r}
source("gf_partialPlot.R")
gf_partialPlot(
  model = fit_caret_rf,
  df = df,
  x.var = "Log_Total_Amount"
)
```

```{r}
source("gf_partialPlot.R")
gf_partialPlot(
  model = fit_caret_rf,
  df = df,
  x.var = "Session_Duration_Minutes"
)
```

```{r}
source("gf_partialPlot.R")
gf_partialPlot(
  model = fit_caret_rf,
  df = df,
  x.var = "Age"
)
```

```{r}
df$predict_review <- predict(fit_caret_rf, newdata = df)


```

```{r}


ggplot(df, aes(x = Session_Duration_Minutes,  y = predict_review)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```
```{r}
ggplot(df, aes(x = Age,  y = predict_review)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```
```{r}
ggplot(df, aes(x = Log_Total_Amount,  y = predict_review)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Predicted Review vs. Log Total Amount",
       y = "Predicted Review",
       x = "Log Total Amount")
```









